---
title: "IDS investigation worksheet"
author: "by Team-Name: User1, User2, User3, User4 & User5"
date: "`r Sys.Date()`"
output: html_document
---

**Note:** You can use this file as you 'working document' where you can try out various investigation ideas and keep notes about your findings. How you use and structure this file is up to you. It is recommended that you keep notes about what you are investigating and what you find as this will make the process of creating your presentation and report easier. Please note that you _do not_ need to submit this file as part of your group project.



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load-lib, message = FALSE}
library(tidyverse)
library(ggplot2)
library(dplyr)
```


```{r load-data}
#tidying up the data

tidy.data <- weekly_ae_activity_20231008 %>%
  select(-Country, -DepartmentType) %>%
  mutate(WeekEndingDate = ymd(WeekEndingDate)) %>%
  mutate(PercentageOver4HoursEpisode = 
           round((NumberOver4HoursEpisode/NumberOfAttendancesEpisode)*100, digits = 1))

data.HBT.pop <- left_join(tidy.data, HBT_area_names_and_population, by ="HBT")

data.TLnames <- Treatement_Location_names%>%
  rename("TreatmentLocation" = "Postcode of Treatment Location")%>%
  left_join(data.HBT.pop, Treatement_Location_names, by ="TreatmentLocation")

col_order <- c("WeekEndingDate", "HBT", "Area name", "Population", "TreatmentLocation",
               "Name of Treatement Location", "NumberOfAttendancesEpisode",
               "NumberWithin4HoursEpisode",
               "PercentageWithin4HoursEpisode", "NumberOver4HoursEpisode",
               "PercentageOver4HoursEpisode", 
               "NumberOver8HoursEpisode", "PercentageOver8HoursEpisode",
               "NumberOver12HoursEpisode", "PercentageOver12HoursEpisode")

data.order <- data.TLnames[, col_order]

data <- data.order %>%
  rename("Week_Ending_Date" = "WeekEndingDate",
         "Health_Board_Area_Code" = "HBT", 
         "Health_Board_Area_Name" = "Area name", 
         "Area_Population" = "Population", 
         "Treatment_Location_Code" = "TreatmentLocation", 
         "Treatment_Location_Name" = "Name of Treatement Location",
         "Total_Attendees" = "NumberOfAttendancesEpisode", 
         "Attendees_within_4hrs" = "NumberWithin4HoursEpisode", 
         "%_within_4hr" = "PercentageWithin4HoursEpisode", 
         "Attendees_Over_4hrs" = "NumberOver4HoursEpisode",
         "%_Over_4hr" = "PercentageWithin4HoursEpisode",
         "Attendees_Over_8hrs" = "NumberOver8HoursEpisode",
         "%_Over_8hr" = "PercentageOver8HoursEpisode", 
         "Attendees_Over_12hrs" = "NumberOver12HoursEpisode",
         "%_Over_12hr" = "PercentageOver12HoursEpisode")

data #tidied data

#graph of attendees per week over the span of a year, for each year
data_graph_1 <- data %>%
  mutate(Year = year(Week_Ending_Date)) %>%
  mutate(date = yday(Week_Ending_Date))


data_graph_1 %>% 
  group_by(Week_Ending_Date) %>%
  summarise(sumdat=sum(Total_Attendees), Year, date) %>% 
  ggplot(aes(x = date, y = sumdat, colour = as.factor(Year))) +
  geom_line() +
  labs(title = "No. Attendees Over the Year",
       subtitle = "2015 - 2023", 
       x = "Day Number", 
       y = "No. of Attendees")

#grpah 5

#Number of Attendees Per Year in Each Health Board Area
data %>%
  select(Week_Ending_Date, Total_Attendees, Health_Board_Area_Name) %>%
  group_by(Week_Ending_Date, Health_Board_Area_Name) %>%
  summarise(Sum_Total_Attendees= sum(Total_Attendees)) %>%
  ggplot(aes(x = Week_Ending_Date, y = Sum_Total_Attendees, colour = Health_Board_Area_Name)) +
  geom_line()+
  labs(x = "Date",
       y = "Sum of Total Attendees",
       title = "Number of Attendees Per Year in Each Health Board Area")
  
```


