---
title: "Impact Of Covid-19 On A&E's In Scotland"
author: "by the A-Team: Anna Scollay, Alice Bullard, Alice Miller, Annie O'Dell"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.retina = 3, dpi = 300, fig.width = 6, fig.asp = 0.618, out.width = "80%") # for better picture resolution
```


```{r load-lib, include = FALSE}
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(knitr)
library(xaringanthemer)
library(devtools)
library(ggplot2)
library(gridExtra)
library(tidymodels)
library(readxl)
```
# Our Data

birds_data

- Species ID: Unique identifier for species (or species group) linking to the Avibase dataset which provide detailed information on taxonomy, synonymy and distribution for the world's birds.

- Species_Name: Bird species name.

- Family: Family-level each bird species belongs to..

- Order: Order-level each bird family belongs to.

- Total Individuals: Number of individual birds measured for each species.

- Female: Number of measured female birds.

- Male: Number of measured male birds.

- Unknown: Number of birds measured with sex unknown.

- Complete Measures: Number of measured birds with a complete set of morphometric trait measurements.

- Hand-Wing Index: 100*DK/Lw, where DK is Kipp’s distance and Lw is wing length.

- Mass: Average body mass (male and female).

- Habitat: Habitat each bird species inhabits.

- Habitat Density: Density of habitat rated from 1 to 3. 1 = Dense habitats, 2 = Semi-open habitats, 3 = Open habitats. 

- Migration: Level of migratory behavior from 1 to 3. 1 = Sedentary, 2 = Partially migratory, 3 = Migratory.

- Trophic Level: Main diet level groups (Carnivore, Omnivore, Herbivore, Scavenger)

- Trophic Niche: Niche diet level, sub categories of Tropic Level groups.

- Primary Lifestyle: Main lifestyle activity taken up by species.

- Range Size: The total area of the mapped range of the species.

